---
title: "Get started"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This tutorial will show you how to use `mimsy` to calculate dissolved gas concentrations from your raw MIMS signal data. There are three steps to using the `mimsy` package:

1. Correctly format your input data
2. Run the `mimsy` function
3. Save your output data using `mimsy.save`

## Download the package

Install the `mimsy` package and load it into your R environment

```{r mimsyLoad, eval=FALSE}
# Use the devtools package to download mimsy from Github
install.packages("devtools")
library(devtools)
devtools::install_github("michelleckelly/mimsy")

# Load mimsy into your R environment
library(mimsy)
```

## Prepare your input data

`mimsy` requires your input data to be in a .csv file, with data columns and column names that are exactly correct. In addition to the columns that are automatically returned by the Bay Instruments MIMS software, `mimsy` needs columns that define the type of sample ("Standard" or "Sample"), the run group (a set of standards followed by a set of samples defines one group), the collection temperature of the sample, and the date that samples were run.

```{r mimsyTemplate}

```

### Type

### Group

### CollectionTemp

### RunDate

## Calculate your dissolved gas concentrations

Run the `mimsy` function on your correctly formatted .csv file. You will need to specify the barometric pressure (`baromet.press`) while samples were run on the MIMS, and specify the units (`units`). Units must be "atm" (atmospheres), "hPa" (hectopascal), "psi" (pounds per square inch), "bar", or "Torr". 

By default, `mimsy` assumes the salinity of standards is zero, no background corrections are necessary, and the `Time` column (returned by the Bay Instruments MIMS software) is referenced to the user's local time zone setting. See ?mimsy for details on how to specify these arguments. 

```{r mimsyRun, eval = FALSE}
data <- mimsy(file = "example.csv", baromet.press = 981.8, units = "hPa",
              std.temps = c(20.1, 25.3))
```

## Inspect the output

`mimsy` returns a list containing five dataframes:

### $results

### $solubility.Concentrations

### $calibration.Factors

### $calibration.DriftCorrection

### $results.full

## Save your data

