---
title: "Get started"
author: "Michelle C Kelly"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This tutorial will show you how to use `mimsy` to calculate dissolved gas concentrations from your raw MIMS signal data. There are three steps to using the `mimsy` package:

1. Correctly format your input data
2. Run the `mimsy` function
3. Save your output data using `mimsy.save`

## Download the package

Install the `mimsy` package and load it into your R environment

```{r mimsyLoad, eval=FALSE}
# Use the devtools package to download mimsy from Github
install.packages("devtools")
library(devtools)
devtools::install_github("michelleckelly/mimsy")

# Load mimsy into your R environment
library(mimsy)
```

## Preparing the input data

## Run the `mimsy` function

## Inspect the output

## Save your data

