---
title: "Vignette: `mimsy`"
author: "Michelle Catherine Kelly"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette: `mimsy`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

```{r install}
# install mimsy
library(devtools)
devtools::install_github("michelleckelly/mimsy")

# load mimsy
library(mimsy)
```

## Prep for mimsy


## Use mimsy

```{r execute}
mimsy::mimsy(filename = "MIMS_KAWN24HR_9Apr18_8thSt_mimsy.csv", barpress = c(981.2, 979.5),
      barpress_units = "hPa", std.temps =  c(9.81, 12.05))
```


## References

Garcia, H., and L. Gordon (1992), \emph{Oxygen solubility in seawater: Better fitting
equations}, Limnology and Oceanography, 37(6).

Benson, B. B. & Krause, D. (1984). \emph{The concentration and isotopic
fractionation of oxygen dissolved in freshwater and seawater in equilibrium
with the atmosphere.} Limnology and Oceanography, 29(3), 620-632.
doi:10.4319/lo.1984.29.3.0620

Stull, D. R. (1947). \emph{Vapor Pressure of Pure Substances. Organic and
Inorganic Compounds.} Industrial & Engineering Chemistry, 39(4), 517-540.
doi: 10.1021/ie50448a022

Hamme, R. C. & Emerson, S. R. (2004). \emph{The solubility of neon, nitrogen and argon
in distilled water and seawater}, Deep-Sea Research I, 51(11), 1517-1528.
